### 2023-11-10 paper/content.tex
Initial paper revisions

The spatial resolution of the inversion results is one of the key advantages of magnetic microscopy over the classic techniques of paleomagnetism.
Therefore, we present the inversion results spatially so that we can evaluate any patterns in their distribution.
Figure~\ref{complex-synthetic-comparison} shows the spatial locations of the 99 sources that were identified and the differences between the estimated dipole moments and the true values.
The depth and dipole moment amplitude of each source are represented by horizontal and vertical bars, respectively.
These two variables are useful proxies for the strength and spatial extent of the signal of each source,
which can tell us about the limitations in strengths and sizes of the source signal that the technique is able to correctly invert.
Figure~\ref{complex-synthetic-comparison}a shows the absolute value of the angular difference between the true and the estimated dipole moment vectors.
Figure~\ref{complex-synthetic-comparison}b shows the percentage difference between the true and estimated dipole moment magnitudes ($|100 \left( m_{true} - m_{estimated}\right) ~/~ m_{true}|$).
Figure~\ref{complex-synthetic-comparison}c shows the $R^2$ coefficient (Equation~\ref{eq_r2}), which is equivalent to the non-dipolarity parameter of \citet{Fu2020} and represents how well the dipolar model is able to explain the observed data of each source.
Figure~\ref{complex-synthetic-comparison}d shows the SNR (Equation~\ref{eq_snr}), which expresses the power of the observed data over that of the inversion residuals.
High SNR values correspond to small inversion residuals which indicate that a dipolar model was able to explain the observed data.
Consequently, the variation of SNR values is similar to that of the $R^2$ coefficient.
Figure~\ref{complex-synthetic-comparison} shows that large errors in the estimated dipole moment direction are associated with low values of $R^2$ and SNR.
Conversely, the correlation between errors in dipole moment magnititude and $R^2$ and SNR is less pronounced, with some poor magnitude estimates being associated with $R^2$ and SNR indicating a good fit of the dipolar model.
It is also noticeable that the majority of cases where the dipole moment direction error is large are associated with deep and low-amplitude sources that are close to shallower or higher-amplitude sources.
These results indicated that $R^2$ and SNR can be used as selection criteria to discard sources with likely high errors in the estimated dipole moment.

@@

In order to assess the algorithm's ability to work with samples featuring higher concentrations of particles, we conducted multiple scenario tests (with 15 repetitions each) by increasing the concentrations of magnetic grains. Figure XX.A depicts the percentage of particle detection ($100($detected particles / total particles$)$) by the algorithm, with the median of the distribution represented by the red line. The same figure also presents the percentage of windows whose inversion results meet the filtering criteria of R-squared greater than or equal to 0.85 and SNR greater than or equal to 5, with the median represented by the blue line.

As the number of particles increases, there is a consistent decrease in the percentage of source detection, particularly due to the spatial constraints of the algorithm, which require magnetic sources to be sparsely distributed for optimal detection. However, as shown in Figure XX.B, by calculating the angular misfit between the mean direction of the model and the mean direction of the detected windows, it is observed that for the unfiltered windows (red line), the angular misfits tend to increase with the concentration of magnetic grains. Nonetheless, when applying the filtering criteria, this misfit tends to approach zero.

@@

@@

Since the pioneering work of \cite{Egli2000}, many methodologies were proposed for solving the inverse problem o micromagnetic data, and for the purpose of comparison we separate them into two categories based on the main estimated parameter by the inversion procedure.
In the first type of approach, the main goal is usually to estimate average magnetization by inverting the whole sample superficial magnetization data commonly by means of unidirectional problem, uniform directions with non-negative variable dipole moments \citep[e.g.,][]{Weiss2007}, including performance enhancement in the spatial domain \citep[e.g.,][]{Myre2019} or frequency domain  \citep[e.g.,][]{Lima2013}.
This methodology can be used to remarkably estimate the average magnetic direction and the total moment direction with the assumption that the particles were all magnetized in the direction of the same induced field \citep[sIRM and/or NRM in basalts,][]{Weiss2007}.
However, this assumption is not always true when dealing with complex samples (i.e., more than one stable direction), which leads to the same drawbacks as the classic paleomagnetic measurements using bulk samples.
The second type of approach has the goal of finding the individual source magnetic properties, which can be done by either inverting the dipole moment of a single source within a cropped section of an upward continued anomaly map \citep[e.g.,][]{Lima2016, Fu2020} or by the insertion of additional information of the sources’ shape, such as micro-computed tomography (microCT) \citep[e.g.,][]{Fabian2019, DeGroot2018, DeGroot2021, Kosters2023}.
The latter further allows unique estimation of magnetic moment configuration of even higher orders components through spherical harmonics expansion constrained by micromagnetic models \citep[e.g.,][]{CortesOrtuno2021, CortesOrtuno2022}.
Such outstanding techniques come with some troubles of having to mechanically select the data for inversion or dealing with the weaknesses of the additional method used.
The MicroCT, for example, is a popular non-destructive technique for high-resolution imaging of the material internal structures, and yet, it is accompanied by some limitations when it comes to paleomagnetic studies: firstly, the technique has a spatial resolution on the order of micrometers, which is not sufficient to directly image the fined grained single-domain magnetite \citep{DeGroot2018}.
The microCT also struggles to discern ferromagnetic (\textit{l.s.}) from non-magnetic/antiferromagnetic minerals (although assigning to them negligible magnetization), as pointed out by \cite{DeGroot2021}, since they usually have similar densities and therefore similar X-ray attenuation \citep{Cnudde2013}.
In any case, the major limitation of microCT lies in the trade-off between the resolution and the sample size, requiring small sample volumes to achieve higher resolutions causing the technique to be too time-consuming.

LINK THE NEXT PARAGRAPH

\subsection{NEW TITLE} %A critical examination of the source detection

@@

The complex synthetic data allows better observation of the strengths and limitations of the windows approach. The main strengths that can be mentioned are: 
(i) the applied technique not only detects most of the modeled sources but also (ii) most of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a).
(iii) The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d).
(iv) Shallow particles grouped in clusters are usually well individualized during window selection, as well as (v) some isolated particles that produce a weak magnetic signal.
% \begin{enumerate}[label=\roman*)]
%   \item The applied technique detects most of the modeled sources;
%   \item The majority of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Figure~\ref{complex-synthetic-comparison}a);
%   \item The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Figure~\ref{complex-synthetic-comparison}b) when $R^2$ and SNR scores are considered high ($\geq 0.85$ and $\geq 5$, respectively) (Figure~\ref{complex-synthetic-comparison}c-d);
%   \item Shallow particles grouped in clusters are usually well individualized during window selection; as well as
%   \item Some isolated particles that produce a weak magnetic signal.
% \end{enumerate}

It is clear from the error bars in Figure~\ref{complex-synthetic-comparison} that deep-seated sources and/or particles with small dipole moments generate worse results, essentially because they will produce weaker anomalous fields in the magnetic maps.
Note nonetheless, that even sources with small dipole moments when close to the surface are adequately modeled by our method, because of the trade-off between signal and noise for grains near the sensor. So we can point out the major limitations observed were: 
(i) the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters.
(ii) The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
(iii) In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values).
As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.

% \begin{enumerate}[label=\roman*)]
%   \item  the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters;
%   \item  The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both.
%   \item In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion, especially the estimated dipole moment (cluster in the Figure~\ref{complex-synthetic-comparison}b with the highest misfit values). As expected, there is a direct relationship between the dipole intensity and depth with the observed errors.
% \end{enumerate}

@@

This remark is also better observed in our simple synthetic sample where the estimated horizontal positions slightly deviate from the true values, which implies small misfit values in the recovered magnetic directions.
Although the estimated magnetic moment for the said sample is satisfactory, this magnetic parameter is more affected by the variations in the depth of the source, which is probably caused by ambiguities.
In summary, in order to estimate all magnetic parameters as precisely as possible the ED must be executed within a data window containing the lowest amount possible of noise since that high-frequency noise sensibility is one of the ED's main limitations.

@@

Our approach relies in the premise of assuming the magnetic anomaly within
the data window is a response of a dipolar source.
The latter is true when working
with particle signals in the SD magnetic domain state since they are uniform magnetized particles with strong dipolar anomalies.
However, \cite{Nagy2017} reported that particles within the PSD domain can record the paleomagnetic field for longer (than SD ones) periods of time, being the stabler and also with strongly non-dipolar characteristics.
Therefore the application of the proposed algorithm to natural samples should fail for domains bigger than PSD state. In these instances, where sources may exhibit higher-order magnetic anomalies influence such as quadrupole or octupole \citep[e.g.,][]{CortesOrtuno2021, Koster2023lava}. In such cases, the dipole model fitting process may not yield satisfactory results. Consequently, the presence of higher-order anomalies within the magnetic sources can impact the accuracy of our inversion results.

@@

Despite the excellent signal-to-noise ratio that the SMM provided with the
proximity of the sensor to the sample, it is worth mentioning that the
measurement noise can still overshadow the responses of very weak/small, or
deep particles, generating unreliable inversion results.
Therefore, it is necessary to keep a check to determine if the inversion reached a satisfactory prediction, such as the coefficient of determination and the
signal-to-noise ratio suggested by \cite{CortesOrtuno2021}.

>>>>

The spatial resolution of the inversion results is one of the key advantages of magnetic microscopy over the classic techniques of paleomagnetism.
Therefore, we present the inversion results spatially so that we can evaluate any patterns in their distribution.
Figure~\ref{complex-synthetic-comparison} shows the spatial locations of the 99 sources that were identified and the differences between the estimated dipole moments and the true values.
The depth and dipole moment amplitude of each source are represented by horizontal and vertical bars, respectively.
These two variables are useful proxies for the strength and spatial extent of the signal of each source,
which can tell us about the limitations in strengths and sizes of the source signal that the technique is able to correctly invert.
Figure~\ref{complex-synthetic-comparison}a shows the absolute value of the angular difference between the true and the estimated dipole moment vectors.
Figure~\ref{complex-synthetic-comparison}b shows the percentage difference between the true and estimated dipole moment magnitudes ($|100 \left( m_{true} - m_{estimated}\right) ~/~ m_{true}|$).
Figure~\ref{complex-synthetic-comparison}c shows the $R^2$ coefficient (Equation~\ref{eq_r2}), which is equivalent to the non-dipolarity parameter of \citet{Fu2020} and represents how well the dipolar model is able to explain the observed data of each source.
Figure~\ref{complex-synthetic-comparison}d shows the SNR (Equation~\ref{eq_snr}), which expresses the power of the observed data over that of the inversion residuals.
High SNR values correspond to small inversion residuals which indicate that a dipolar model was able to explain the observed data.
Consequently, the variation of SNR values is similar to that of the $R^2$ coefficient.
Figure~\ref{complex-synthetic-comparison} shows that large errors in the estimated dipole moment direction are associated with low values of $R^2$ and SNR.
Conversely, the correlation between errors in dipole moment magnitude and $R^2$ and SNR is less pronounced, with some poor magnitude estimates being associated with $R^2$ and SNR indicating a reasonable fit by dipolar model.
It is also noticeable that the majority of cases where the direction misfit is given by deep and low-amplitude sources that are close to shallower or higher-amplitude sources.
These results indicated that $R^2$ and SNR can be used as selection criteria to discard sources with likely high errors in the estimated dipole moment.

@@

In order to assess the algorithm's ability to work with samples featuring higher concentrations of particles, we conducted multiple scenario tests (with 15 repetitions each) by increasing the concentrations of magnetic grains. Figure~\ref{grain-concentration}a depicts the percentage of particle detection by the algorithm, with the median of the distribution represented by the red line. The same figure also presents the percentage of windows whose inversion results meet the filtering criteria of $R^2 \geq 0.85$ and $SNR \geq 5$, with the median represented by the blue line. As the number of particles increases, there is a consistent decrease in the percentage of source detection, particularly due to the spatial constraints of the algorithm, which require magnetic sources to be sparsely distributed for optimal detection. However, as shown in Figure~\ref{grain-concentration}b, by calculating the angular misfit between the vector sum direction of the model and the vector sum direction of the detected windows, it is observed that for the unfiltered windows (red line), the angular misfits tend to increase with the concentration of magnetic grains. Nonetheless, when applying the filtering criteria, this misfit tends to approach zero.

@@

\begin{figure}[tb!]
  \centering
  \includegraphics[width=1\linewidth]{paper/figures/grain-concentration.png}
  \caption{ 
  Different grains simulations randomly replicated (N=15) to assess the effectiveness of the algorithm to detect the magnetic sources. a) The efficacy of the detection was measured by percentage of sources identified ($100 \left( N_{detected~sources} ~/~ N_{true~sources} \right)$) and showed as box-plots for both unfiltered detection (median represented as the red line) and filtered detection ($R^2 \geq 0.85$ and $SNR \geq 5$, blue line). b) The same representation is displayed for the angular misfit between the true vector sum and both results for the filtered and unfiltered sum of vectors detected. The blue/red dots represent outliers in the distributions.
  }
  \label{grain-concentration}
\end{figure}

@@

Since the pioneering work of \cite{Egli2000}, many methodologies were proposed for solving the inverse problem o micromagnetic data, and for the purpose of comparison we separate them into two categories based on the main estimated parameter by the inversion procedure.
In the first type of approach, the main goal is usually to estimate average magnetization by inverting the whole sample superficial magnetization data commonly by means of unidirectional problem, uniform directions with non-negative variable dipole moments \citep[e.g.,][]{Weiss2007}, including performance enhancement in the spatial domain \citep[e.g.,][]{Myre2019} or frequency domain  \citep[e.g.,][]{Lima2013}.
This methodology can be used to remarkably estimate the average magnetic direction and the total moment direction with the assumption that the particles were all magnetized in the direction of the same induced field \citep[sIRM and/or NRM in basalts,][]{Weiss2007}.
However, this assumption is not always true when dealing with complex samples (i.e., more than one stable direction), which leads to the same drawbacks as the classic paleomagnetic measurements using bulk samples.
The second type of approach has the goal of finding the individual source magnetic properties, which can be done by either inverting the dipole moment of a single source within a cropped section of an upward continued anomaly map \citep[e.g.,][]{Lima2016, Fu2020} or by the insertion of additional information of the sources’ shape, such as micro-computed tomography (microCT) \citep[e.g.,][]{Fabian2019, DeGroot2018, DeGroot2021, Koster2023}.
The latter further allows unique estimation of magnetic moment configuration of even higher orders components through spherical harmonics expansion constrained by micromagnetic models \citep[e.g.,][]{CortesOrtuno2021, CortesOrtuno2022}.
Such outstanding techniques come with some troubles of having to mechanically select the data for inversion or dealing with the weaknesses of the additional method used.
The MicroCT, for example, is a popular non-destructive technique for high-resolution imaging of the material internal structures, and yet, it is accompanied by some limitations when it comes to paleomagnetic studies: firstly, the technique has a spatial resolution on the order of $\mu m$, which is not sufficient to directly image the fined grained single-domain magnetite \citep{DeGroot2018}.
The microCT also struggles to discern ferromagnetic (\textit{l.s.}) from non-magnetic/antiferromagnetic minerals (although assigning to them negligible magnetization), as pointed out by \cite{DeGroot2021}, since they usually have similar densities and therefore similar X-ray attenuation \citep{Cnudde2013}.
In any case, the major limitation of microCT lies in the trade-off between the resolution and the sample size, requiring small sample volumes to achieve higher resolutions causing the technique to be too time-consuming.

\subsection{TGA-guided windows: magnetic source detection using LoG Kernel} %A critical examination of the source detection

@@

Another main limitation arises from the spatial constraints inherent to our approach. Specifically, the blob detection struggles when dealing with sources that are closely spaced or overlapping, in the last case, it is unable to discern the magnetic sources. Through multiple tests (Figure~\ref{detection-map}), we have found that, for shallow sources, with sensor-sample height $\leq 5 \mu m$, a minimum separation distance of approximately 14 $\mu m$ is necessary to achieve successful source detection. This distance steadily increases for deeper seated sources, requiring around 20 $\mu m$ of separation for sensor-sample height of 25 $\mu m$. This immensely impacts the algorithm's detection capacity, as the particle count escalates, there is a persistent decline in the rate of successful source detection, as shown in Figure~\ref{grain-concentration}a. In essence, the algorithm's effectiveness diminishes as sources become more densely packed, posing a challenge in scenarios where particle density is high (e.g., extrusive rocks). On the other hand, as shown in our complex synthetic and real rock samples the technique could be densely applied to rocks with a lower concentration of magnetic grains (e.g., carbonates), due to its high successful rate for detection of sources.

\begin{figure}[tb!]
\centering
\includegraphics[width=0.5\linewidth]{paper/figures/detection-map.png}
\caption{
Simulation to assess the algorithm's ability to discern between two sources, with different horizontal separation and varying sensor-sample heights. The black line represents the boundary where the algorithm's is capable of identify two sources.
}
\label{detection-map}
\end{figure}

@@

This remark is also better observed in our simple synthetic sample where the estimated horizontal positions slightly deviate from the true values, which implies small misfit values in the recovered magnetic directions.
Although the estimated magnetic moment for the said sample is satisfactory, this magnetic parameter is more affected by the variations in the depth of the source, which is probably caused by ambiguities. Despite the fact that we use the ED original formulation dedicated to dipolar sources ($\eta = 3$), the technique did not present significant difficulties in estimating the 3D position of non-dipolar sources. This adaptability allows for an extension of its application to magnetic particles with a strong influence of higher-order magnetic anomalies. Lastly, as shown in the complex synthetic sample, the ED can reliably estimated the position of well-individualized sources, however when dealing with clustered particles the Euler's solutions start to deviate from the true positions. This deviation is particularly notable in the z-coordinate, which is showed by \cite{Oliveira2015Estimation} to not affect the magnetization direction estimation. However, the scenarios with closely spaced magnetic sources, where the inherent assumptions of Euler's method encounter challenges in accurately capturing the complex interactions of magnetic field anomaly caused by each nearby source, greatly interfering in the dipole intensity estimation.

In summary, in order to estimate the Cartesian position as precisely as possible the ED must be executed within a data window containing only one magnetic source and the lowest amount possible of noise, since the high-frequency noise sensibility is also one of the ED's limitations. As these assumptions deviate from the ideal case, the ED estimates also become less accurate.

@@

Our approach relies in the premise of assuming the magnetic anomaly within
the data window is a response of a dipolar source.
The latter is true when working
with particle signals in the SD magnetic domain state since they are uniform magnetized particles \citep{Dunlop1997} with strong dipolar anomalies.
However, \cite{Nagy2017} reported that particles within the PSD domain can record the paleomagnetic field for longer (than SD ones) periods of time, being the stabler and also with strongly non-dipolar characteristics.
Hence, the implementation of the proposed algorithm in natural samples is expected to be unsuccessful in domains other than the SD state. In these instances, where sources may exhibit higher-order magnetic anomalies influence, such as quadrupole or octupole \citep[e.g.,][]{CortesOrtuno2021}, the dipole model fitting process may not yield satisfactory results. Consequently, the presence of higher-order anomalies within the magnetic sources can impact the accuracy of our inversion results.

@@

Despite the excellent signal-to-noise ratio that the SMM provided with the
proximity of the sensor to the sample, it is worth mentioning that the
measurement noise can still overshadow the responses of very weak/small, or
deep particles, generating unreliable inversion results.
Therefore, it is necessary to keep a check to determine if the inversion reached a satisfactory prediction, such as the coefficient of determination \citep[dipolarity test, ][]{Fu2020} and the
signal-to-noise ratio suggested by \cite{CortesOrtuno2021}. This filtering process is better observed in the complex synthetic data where it becomes evident the direct relationship between the poorly estimated dipole moment and direction with the lowest values for both $R^2$ and SNR scores (Figure~\ref{complex-synthetic-comparison}).
